\documentclass[../mattg_ti-fi_lit-review.tex]{subfiles}

\subsection{Theory}

\subsection{Spin Torque}
\subsubsection{Torque}
\subsubsection{Spin Torque Angle - Spin Torque Efficiency}
\subsubsection{Magnetization}
Larmor Equation:
\begin{align}
	\vec{T} = \mu_0 \vec{\textbf{M}}\times \vec{\textbf{H}}
\end{align}
Torque:
\begin{align}
	\frac{d\vec{L}}{dt}=\vec{T}
\end{align}
Gyromagnetic Ratio
\begin{align}
	\gamma\equiv\frac{\mu}{L}
\end{align}
Magnetization Change
\begin{align}
\frac{d\vec{M}}{dt} = \gamma\vec{T}
\end{align}
\subsubsection{Critical Current Density}
\subsubsection{Charge Spin Conversion Efficiency}
\subsection{Experiments}
\textcolor{red}{TODO: go through past experiments. Take significant of results, sample fabrication, experimental methodology}.
\subsubsection{Conductive films (non-insulating)}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
	\item (2012) Liu et al. Beta-Ta with CoFeB  (Observed Torque, 2.1mT for 0.7mA, Sputtered/evaporated onto surfaces) 
	\item (Apr 2013) Chang et al. Observation of QAHE - Cr0.15(Bi0.1Sb0.9)1.85Te3  - 5 quintuple layers, on SrTiO3, grown through MBE. Mobilities lower than 1000cm2/Vs. Chromium doping to make ferromagnetic, RHEED, ARPES.
	\item (Apr 2014) Y. Fan et al. Epitaxial (Bi0.5Sb0.5)2Te3/(Cr0.08Bi0.54Sb0.38)2Te3 bilayer films, 3 \& 6 Quintuple layers respectively. AHE. ($\theta$SH reported to be very large??). Look again at device fab. 
	\item (2014) A.R. Mellnik et al.  8nm Bi2Se3 and 8nm / 16nm Ni81Fe19 Bilayer (MBE \& Sputtering) (Observed Torque Applied, Torque per unit moment Tparallel = 2.7 x 10-5, T¬perp=3.7 x 10-5), 
	$\omega$SH\_parallel=(2.0 ± 0.4) × 105/2e $\omega$-1 m-1 and  $\sigma$SH\_perp  = (1.6 ± 0.2) × 105/2e $\omega$-1 m-1
	\item  (2016) P. Li et al. Pt/BaFe12O19 – Torque observed, value not mentioned. 
 	\item (Aug 2017) J. Han et al. (MIT \& Samarth Grp) - GaAs / Bi2Se3(7.4nm)/CoTb(4.6nm)/SiNx(3 nm) - Magnetron and MBE. Critical current density 2.8 x 106 A/cm$^2$. 
	0.16 effective SH angle. Large discrepancies noted in this paper. Also searched (Bi, Sb)2Te3, and compared to Pt and other heavy metals.
	\item (Sep 2017) K. Yasuda et al. Crx(Bi1-ySby)2-xTe3/(Bi1-ySby)2Te3 - Magnetic non-magnetic heterostructures. Second Harmonic Voltage goverened by asymmetric magnon scattering gives incorrect value of CSC (Charge spin conversion efficiency).
	\item \textbf{(Nov 2017) Y. Wang et al. Bi2Se3/NiFe Heterostructure, Describes contamination of BS, 2DEG to the SOT effects from TSS affecting Bi2Se3 systems. Also used Bi2Se3/Co40Fe40B20, and Bi2Se3 (8 QL)/Py(6nm). Growths on Al2O3. Used MgO and Al2O3 to protect. Also SiO2 used for Py devices. Iron Milling used with Photolithography.}
	\item (Jul 2018) Khang et al. Bi0.9Sb0.1, 5/10nm Mn0.45Ga0.55 3nm layers\\ $\sigma\approx2.5\times10^5\omega^{-1}m^{-1},\theta_{SH}\approx 52$ and $\theta_{SH}\approx1.3\times10^7\hbar / 2e \omega^{-1}m^{-1}$
	\item (Jul 2018) Mahendra DC et al. BixSe(1-x)/Co20Fe60B20 (Magentron Sputtered)\\
	$\theta_S = 18.62$ (d.c. planar Hall) \&8.67 (ST-FMR) [Si/SiO2/MgO (2nm)/BixSe(1-x)(tBS nm)/ CoFeB(5nm)/MgO(2nm)/Ta(5nm)] tBS=4,6,8,16,40nm
	\item (Jan 2018) Y. Li et al.  Epitaxial (001) SmB6/Si thin films (50nm +) with MgO(1.5)/CFB(Co40Fe40B20, 1)/$\beta-W(0.8)$ – DC Magnetron sputtering / DC\&RF magnetron…
\end{itemize}
\subsubsection{Insulating Films}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
	\item (2019) P. Li et al. Bi2Se3 and FI - BaFe12O19 \url{https://advances.sciencemag.org/content/5/8/eaaw3415}
	\item YIG (2016) H. Wang et al. \url{https://journals-aps-org.ezproxy.lib.monash.edu.au/prl/pdf/10.1103/PhysRevLett.117.076601}
\end{itemize}

\subsection{Non-TI Devices}
- (2010) Co layer 0.6nm, with Pt 3nm and AlO\_x layers 2nm, transverse magnetic fields. https://www-nature-com.ezproxy.lib.monash.edu.au/articles/nmat2613?proof=trueMay

